Turbulence and Particle Acceleration in Giant Radio Halos: the Origin of Seed Electrons

### 2016-05-30 paper.tex
fixing a few typos

All of the physics of turbulent reacceleration is effectively encapsulated in the diffusion coefficient \citep{brunetti07}, which can be rewritten as \citep{miniati15}:
\begin{equation}
D_{\rm pp}(p) = \frac{ p^{2} \pi I_{\theta}(c_{s}/c)}{8 c} \langle k \rangle_{\mathcal{W}} \langle (\delta v_{\rm c})^{2} \rangle
\label{eqn:diffusion}
\end{equation}
where $I_{\theta}$ averages interaction rates over the CR pitch angle $\theta$; $I_{\theta} \approx 5$ for ICM conditions, and
\begin{equation}
\langle k \rangle_{\mathcal{W}} = \frac{1}{\langle (\delta \varv_{\rm c})^{2} \rangle} \int_{k_{\rm L}}^{k_{\rm cut}} dk \, k \, \mathcal{W}(k) \approx \frac{s-1}{2-s} k_{\rm L} \left( \frac{k_{\rm cut}}{k_{\rm L}} \right)^{2-s} 
\label{eqn:k_W} 
\end{equation}
is an energy-averaged wavenumber, $k_{\rm L}, k_{\rm cut}$ are the wavenumbers associated with the outer scale and the cutoff scale respectively, and we have assumed a total energy spectrum (both kinetic and potential energy, where the two are assumed to be in equipartition \citep{sarkar11}): 
\begin{equation}
\mathcal{W}(k) = \frac{(s-1) \langle (\delta \varv_{\rm c})^{2} \rangle}{k_{\rm L}} \left( \frac{k}{k_{\rm L}} \right)^{-s} 
\label{eqn:Wk}
\end{equation}
which defines the normalization $\langle (\delta \varv_{\rm c})^{2} \rangle$ (the subscript 'c' emphasizes that we specialize to compressive modes). Intuitively, we can understand the form of the the diffusion coefficient from the fact that for second-order Fermi acceleration, $\dot{p} \sim \Delta p/\tau \sim p\,\langle k c \rangle_{\mathcal{W}} (\delta \varv_{\rm c}/c)^{2}$, where $\tau^{-1} \sim \langle k c \rangle_{\mathcal{W}}$ is the energy averaged wave-particle interaction rate, and $\Delta p \sim p (\delta \varv_{\rm c}/c)^{2}$ is the typical momentum change during wave particle scattering. Thus: 
\begin{equation}
D_{\rm pp} \sim p \dot{p} \sim p^{2} \langle k c \rangle_{\mathcal{W}} \left( \frac{\delta \varv_{\rm c}}{c} \right)^{2}. 
\end{equation}

Equations \ref{eqn:diffusion} and \ref{eqn:Wk} make the important aspects of turbulence clear: the energy in compressive modes $\langle (\delta v_{\rm c})^{2} \rangle$, the inner and outer scale ($k_{\rm cut}$ and $k_{\rm L}$), and the slope of the energy spectrum $s$. All of these can vary spatially and temporally.

We adopt a Kraichnan spectrum ($s=3/2$) for the fast modes, as seen in simulations \citep{cho03}. This can also be written as: 
\begin{equation}
  \label{eq:Wk}
  \mathcal{W}(k) =
\sqrt{2/7\,I_0\,\rho\,\langle \Vph \rangle}\,k^{-3/2},
\end{equation}
where $I{\rm 0}$ is the volumetric injection rate of turbulence (which is assumed to be constant), $v_{\rm ph}\approx c_{\rm s}$ is the phase speed of waves. It is important to realize that this only applies at small scales, which turbulence is MHD. At large scales, turbulent motions are subsonic (${\mathcal M}_{\rm s} \sim 0.2-0.5$) but super-Alfv{\'e}nic (${\mathcal M}_{\rm A} \sim 5$). Thus, except in the case where motions are transonic and weak shocks become important, motions are fundamentally hydrodynamic and turbulence follows a Kolmogorov ($s=5/3$) spectrum. However, while the turbulent energy density $U_{\rm t} \propto l^{2/3}$ decreases at small scales, the magnetic energy density is scale-independent. Thus, at some scale $l_{\rm A} \sim l_{\rm o} {\mathcal M}_{\rm A}^{-3}$, where $U_{\rm turb} \sim U_{B}$, the magnetic field becomes dynamically important, and turbulence transitions to the MHD regime. The cutoff scale in equation \ref{eqn:Wk} can be computed by setting the turbulent cascade rate for fast modes to the transit time damping rate on thermal electrons. This yields \citep{brunetti07,miniati15}:
\begin{equation}
k_{\rm c} \approx A k_{\rm outer} \frac{\langle v_{\rm outer}^{2}\rangle^{2}}{c_{\rm s}^{4}}
\label{eqn:k_c} 
\end{equation}
where $A \approx 11 000$. This gives $2\upi/k_\rmn{cut}\sim 0.1-1$~kpc in the ICM. This constitutes an
effective mean free path for CRs, unless plasma instabilities can
mediate interactions between turbulence and particles on smaller
scales \citep{brunetti11}, a possibility we discuss in \S\ref{sect:self-limiting}. Another possibility is that compressible modes dissipate in weak shocks, resulting in Burgers' turbulence $s=2$ \citep{kowal10,porter15,miniati15}. If Burgers turbulence dominates, then particle acceleration rates are too slow in the face of cooling processes to explain radio halos \citep{miniati15}, and an alternative model for radio halos is required.

@@

The important effects are best summarized in terms of the acceleration time: 
\begin{equation}
\tau_{\rm D}= \frac{p^{2}}{4 D_{\rm pp}}
\end{equation}
which is momentum independent. This should be compared against the lifetime of turbulence, $\tau_{\rm cl}$, and the cooling time $\tau_{\rm cool}$. In Table~\ref{tab:timescales}, we show both the thermal quantities and the
timescales for CR cooling and (re)acceleration for three different
spatial regions of the RH. The densities \citep{1992A&A...259L..31B} and temperatures \citep{2009ApJ...696.1886B,2001A&A...365L..67A} are derived from X-ray observations. To calculate synchrotron cooling times, we use B-fields derived from Faraday rotation measurements \citep{bonafede10}. To calculate the acceleration time, we need to assume an outer scale. We assume an injection scale $k_{\rm L}=2\pi/\lambda_{\rm L}$, where $\lambda_{\rm L} = 100$ kpc, which were assumptions adopted in previous work \citep{2006MNRAS.366.1437S,brunetti07, brunetti11}. This lengthscale corresponds to an eddy turnover time on the outer scale of $v_{L} k_{\rm L}^{-1} \sim 1.2$Gyr if $v_{\rm L} \sim 500 {\rm km \, s^{-1}}$, as is characteristic of a merger. Note that hydrodynamical simulations of clusters have sometimes found larger $\lambda_{\rm L}$, in some cases comparable to the size of the cluster \citep{vazza11,miniati15}. This choice is degenerate with $X_{\rm tu}$; in \S\ref{sect:self-limiting} we argue that $k_{\rm A}$ is a more appropriate choice, but also find that when the decay time is appropriately scaled, we are relatively insensitive to the choice of $k_{\rm L}$. We present $\tau_{\rm D}$ for 3 different models, which we discuss in the next section. The reacceleration timescale
$\tau_{\rm D}$ is similar between our three models, where the
difference comes from turbulent profile parameterized with
$\alpha_\rmn{tu}$. This implies that even small differences in the
turbulent profile could impact the seed CRs significantly. Finally, we adopt an duration of acceleration $\tau_{\rm cl} = 650$ Myr, in line with previous assumptions \citep{brunetti07}, roughly corresponding to the turbulent decay time. 
\begin{table}
  \caption{Thermal quantities and timescales for different spatial
    regions in a Coma like cluster.}
\begin{tabular}{l c  c c}
\hline
\hline
& & spatial regions & \\
 & $0.1\,\RH^{(2)}$ & $0.3\,\RH^{(2)}$ & $\RH^{(2)}$   \\
\hline
thermal quantities$^{(1)}$ & & & \\
$\rho$ [$10^{-27}$ g cm$^{-3}$] & 3.0 & 1.6 & 0.15 \\
$T$ [$10^{8}$ K] & 1.4 & 1.0 & 0.58 \\
\hline
timescales$^{(1)}$ & & & \\
$\tau_\rmn{D}$(\Mprimary)$^{(3)}$ [Gyr] & 0.45 & 0.44 & 0.39 \\
$\tau_\rmn{D}$(\Mstream)$^{(3)}$  [Gyr] & 0.50 & 0.47 & 0.34 \\
$\tau_\rmn{D}$(\Mflatturb)$^{(3)}$  [Gyr] & 0.69 & 0.56 & 0.27 \\
$\tau_\rmn{IC/sync}(P=10^4\,m_ec)^{(4)}$ [Gyr] & 0.11 & 0.15 & 0.22 \\
$\tau_\rmn{had}(P=100\,m_pc)^{(5)}$ [Gyr] & 2.4 & 4.5 & 47 \\
$\tau_\rmn{Coul}(P=\,m_ec)^{(6)}$  [Gyr] & 0.0092 & 0.017 & 0.17 \\
\hline
\end{tabular}
\begin{quote}
 Notes: \\ 
 (1) Median quantities from our simulated post-merging cluster g72a derived 
 during last 300~Myrs in time. \\
 (2) Radius of the giant radio halo in Coma where $\RH\approx0.6\,R_{200}$.\\
 (3) Fermi-II reacceleration for both electrons and protons at all energies.\\
 (4) Inverse Compton and synchrotron cooling for electrons.\\
 (5) Catastrophic losses for protons.\\
 (6) Coulomb cooling for electrons (protons factor $m_e/m_p$ smaller).\\
 \label{tab:timescales}
  \end{quote}
\end{table}

@@

We have seen that the radio luminosity depends very sensitively on $\tau_{\rm cl}/\tau_{\rm D}$, through the very sensitive dependence on $X_{\rm tu}$ (Fig \ref{fig:param_comp}; note from equation \ref{eq:Dpp_scaling} that $\tau_{\rm D} \propto X_{\rm tu}^{-2}$). Since this raises questions of fine-tuning in $X_{\rm tu}$ to explain the observations, it is worth understanding in more detail. Let us ignore cooling. This is a good approximation for the CR protons, for instance, since hadronic cooling times are long\footnote{In principle, CRp can still lose energy via wave heating, at a rate $v_{\rm A} \cdot \nabla P_{c}$, but we assume CR streaming is suppressed during mergers, when they are spatially confined by scattering.} In this case, $\dot{p}= p/\tau_{\rm D}$, and after a time $\tau_{\rm cl}$, we have $p \rightarrow p \, {\rm exp}(\tau_{\rm cl}/\tau_{\rm D})$ (where we have used the fact that $\tau{\rm D}$ is momentum independent). For an initial power law distribution function $f(p) = C p^{-\alpha}$, this momentum increase can be rewritten as a change of normalization, $f(p) = \tilde{C} p^{-\alpha}$, where $\tilde{C} = {\rm exp}( \alpha \tau_{\rm cl}/\tau_{\rm D})$. A slightly more careful derivation by direct solution of the Fokker-Planck equation yields: 
\begin{equation}
  \dot{f} = {{\partial }\over{\partial p}}
  p^2 D_{\rm pp} {{\partial }\over{\partial p}} \frac{f}{p^2}\,,
\end{equation}
with the analytic solution given by
\begin{equation}
  \label{eq:Creacc}
  \tilde{C}= 
  C \exp{\left[\frac{(2+\alpha)(\alpha-1)}{4}\frac{\tau_\rmn{cl}}{\tau_\rmn{D}}\right]}\,.
\end{equation}
This exponential sensitivity to $\tau_{\rm cl}/\tau_{\rm D}$ lies behind the extreme sensitivity to $X_{\rm tu}$ in Fig \ref{fig:param_comp}.

@@

An obvious candidate for $\tau_{\rm decay}$ is: 
\begin{equation}
\tau_{\rm edd} = \frac{L}{v_{\rm c}}
\end{equation}
i.e., the eddy turnover time at the outer scale. This is subject to uncertainties about the location of the outer scale $L$; estimates in the literature range from $L \sim 0.1-1$ Mpc. It is also worth remembering that MHD turbulence only applies below $l < l_{\rm A}$, where $l_{\rm A}$ is the Alfv{\'e}n scale where $v \sim v_{\rm A}$. Invoking fast modes, Kraichnan scalings, etc, is only valid below these scales. For $l > l_{\rm A}$, turbulence is basically hydrodynamic and Kolmogorov in nature. The momentum diffusion coefficient can thus be split up into two components: 
\begin{equation}
D_{\rm pp} = D_{\rm pp}^{\rm HD}[l_{\rm A} < l < L] + D_{\rm pp}^{\rm MHD}[l_{\rm c} < l < l_{\rm A}]
\end{equation}
From equations (\ref{eqn:diffusion}) and (\ref{eqn:k_W}), and assuming Kolmogorov turbulence from $l_{\rm A} < l < L$ and Kraichnan turbulence from $l_{\rm c} < l < l_{\rm A}$, we get: 
\begin{equation}
\frac{D_{\rm pp}^{\rm HD}}{D_{\rm pp}^{\rm MHD}} = \frac{k_{\rm L}^{2/3}k_{\rm A}^{1/3} v_{\rm c}^{2}}{k_{\rm A}^{1/2}k_{\rm c}^{1/2} v_{\rm A}^{2}} = A^{-1/2} \beta = 0.5 \left( \frac{\beta}{50} \right) 
\end{equation}
where we have used $k_{\rm A} = k_{\rm L} M_{A}^{3}$ (from Kolmogorov scalings, and $M_{\rm A} = v_{\rm c}/v_{\rm A}$ is the Alfv{\'e}n Mach number of the turbulence, $\beta \approx c_{\rm s}^{2}/v_{\rm A}^{2}$, and the cutoff scale $k_{\rm c}$ (equation \ref{eqn:k_c}).  Thus, the contribution of the two regimes are comparable (as we shall see later, this is not true if one incorrectly assumes Kraichnan scalings apply from the outer scale $L$; this boosts the contribution of the scales where $l_{\rm A} < l < L$ by a factor $M_{\rm A}^{1/2}$). 
Henceforth, we can consider the outer scale of the fast modes to be $l_{\rm A}$, with velocity $v_{\rm A}$. The turbulent reacceleration time is: 
\begin{equation}
\tau_{\rm D} = \frac{p^{2}}{4 D_{\rm pp}} = C_{\rm D} A^{1/2} c k_{\rm A} \frac{v_{A}^{2}}{c_{\rm S}^{2}}
\label{eqn:tau_D} 
\end{equation}
where $C_{\rm D} = 2/(5 \pi)$, $A\approx 11000$, and we have used equation \ref{eqn:k_c}. The turbulence decay time is given by the cascade time of fast modes (\citet{2004ApJ...614..757Y}; see equation \ref{eqn:cascade}): 
\begin{equation}
\tau_{\rm decay} = \frac{c_{\rm s}}{v_{\rm A}^{2} k_{\rm A}}\,.
\label{eqn:tau_decay} 
\end{equation} 
This can be related to the eddy turnover time at the outer scale $\tau_{\rm edd}$ from $l_{\rm A} = L M_{\rm A}^{-3}$ to: 
\begin{equation}
\tau_{\rm decay} =\frac{L}{v_{\rm c}} \left(\frac{c_{\rm s}}{v_{\rm c}} \right)^{2} \frac{v_{\rm A}}{c_{\rm s}} = 0.7 \, \tau_{\rm edd} \left(\frac{\tilde{X}_{\rm tu}}{0.2}\right)^{-1}  \left(\frac{\beta}{100} \right)^{-1/2}
\end{equation}
where we have used $\tilde{X}_{\rm tu} \approx (v_{\rm c}/c_{\rm s})^{2}$ and $c_{\rm s}/v_{\rm A} = \beta^{1/2}$. Thus $\tau_{\rm decay} \sim \tau_{\rm edd}$, but we prefer to use the expression for $\tau_{\rm edd}$, since it is more physically correct and leads to more revealing scalings.

From equations \ref{eqn:tau_D} and \ref{eqn:tau_decay}, we obtain:
\begin{equation}
\frac{\tau_{\rm cl}}{\tau_{\rm D}} \approx \frac{\tau_{\rm decay}}{\tau_{\rm D}} = 0.1 \left( \frac{c_{\rm s}}{1500 \, {\rm km \, s^{-1}}} \right) \left( \frac{\beta}{50} \right)^{-1}\,. 
\label{eqn:ratios_TTD} 
\end{equation}
Remarkably, this expression is independent of properties of the turbulence such as $v_{\rm c},L$, or $l_{\rm A}$, and depends only on properties of the plasma ($c_{\rm s},\beta$). Ultimately, this arises because the timescale on which a fast mode wave transfers energy due to wave particle interactions in second order Fermi acceleration, $\tau_{\rm p} \sim (k_{\rm p} c (v_{\rm c}/c)^{2} )^{-1}$, is closely related to the timescale on which it cascades due to wave-wave interactions, $\tau_{\rm w} \sim c_{\rm s}/(k_{\rm w} v_{\rm c}^{2})$. This implies $\tau_{\rm w}/\tau_{\rm p} \sim (c_{s}/c)(l_{\rm w}/l_{\rm p})$, where $l_{\rm w} \sim l_{\rm A}$ is the outer scale on which the fast mode cascade begins, and $l_{\rm p} \sim (l_{\rm A} l_{\rm c})^{1/2}$ is the characteristic wavelength for wave-particle interactions. For transit time damping and an outer scale of $l_{\rm A}$, we have $l_{\rm c} \sim (m_{\rm e}/m_{\rm p}) l_{\rm A} (c_{\rm s}/v_{\rm A})^{4}$. We thus have $\tau_{\rm w}/\tau_{\rm p} \sim (c_{\rm s}/c)(l_{\rm A}/l_{\rm c})^{1/2} \sim (c_{\rm s}/c) (m_{\rm p}/m_{\rm e})^{1/2} \beta^{-1}$. More careful consideration of dimensionless factors boosts this estimate by an order of magnitude to give equation \ref{eqn:ratios_TTD}.

>>>>

All of the physics of turbulent reacceleration is effectively encapsulated in the diffusion coefficient \citep{brunetti07}, which can be rewritten as \citep{miniati15}:
\begin{equation}
D_{\rm pp}(p) = \frac{ p^{2} \pi I_{\theta}(c_{s}/c)}{8 c} \langle k \rangle_{\mathcal{W}} \langle (\delta \varv_{\rm c})^{2} \rangle
\label{eqn:diffusion}
\end{equation}
where $I_{\theta}$ averages interaction rates over the CR pitch angle $\theta$; $I_{\theta} \approx 5$ for ICM conditions, and
\begin{equation}
\langle k \rangle_{\mathcal{W}} = \frac{1}{\langle (\delta \varv_{\rm c})^{2} \rangle} \int_{k_{\rm L}}^{k_{\rm cut}} dk \, k \, \mathcal{W}(k) \approx \frac{s-1}{2-s} k_{\rm L} \left( \frac{k_{\rm cut}}{k_{\rm L}} \right)^{2-s} 
\label{eqn:k_W} 
\end{equation}
is an energy-averaged wavenumber, $k_{\rm L}, k_{\rm cut}$ are the wavenumbers associated with the outer scale and the cutoff scale respectively, and we have assumed a total energy spectrum (both kinetic and potential energy, where the two are assumed to be in equipartition \citep{sarkar11}): 
\begin{equation}
\mathcal{W}(k) = \frac{(s-1) \langle (\delta \varv_{\rm c})^{2} \rangle}{k_{\rm L}} \left( \frac{k}{k_{\rm L}} \right)^{-s} 
\label{eqn:Wk}
\end{equation}
which defines the normalization $\langle (\delta \varv_{\rm c})^{2} \rangle$ (the subscript 'c' emphasizes that we specialize to compressive modes). Intuitively, we can understand the form of the the diffusion coefficient from the fact that for second-order Fermi acceleration, $\dot{p} \sim \Delta p/\tau \sim p\,\langle k c \rangle_{\mathcal{W}} (\delta \varv_{\rm c}/c)^{2}$, where $\tau^{-1} \sim \langle k c \rangle_{\mathcal{W}}$ is the energy averaged wave-particle interaction rate, and $\Delta p \sim p (\delta \varv_{\rm c}/c)^{2}$ is the typical momentum change during wave particle scattering. Thus: 
\begin{equation}
D_{\rm pp} \sim p \dot{p} \sim p^{2} \langle k c \rangle_{\mathcal{W}} \left( \frac{\delta \varv_{\rm c}}{c} \right)^{2}. 
\end{equation}

Equations \ref{eqn:diffusion} and \ref{eqn:Wk} make the important aspects of turbulence clear: the energy in compressive modes $\langle (\delta \varv_{\rm c})^{2} \rangle$, the inner and outer scale ($k_{\rm cut}$ and $k_{\rm L}$), and the slope of the energy spectrum $s$. All of these can vary spatially and temporally.

We adopt a Kraichnan spectrum ($s=3/2$) for the fast modes, as seen in simulations \citep{cho03}. This can also be written as: 
\begin{equation}
  \label{eq:Wk}
  \mathcal{W}(k) =
\sqrt{2/7\,I_0\,\rho\,\langle \Vph \rangle}\,k^{-3/2},
\end{equation}
where $I_{\rm 0}$ is the volumetric injection rate of turbulence (which is assumed to be constant), $\varv_{\rm ph}\approx c_{\rm s}$ is the phase speed of waves. It is important to realize that this only applies at small scales, which turbulence is MHD. At large scales, turbulent motions are subsonic (${\mathcal M}_{\rm s} \sim 0.2-0.5$) but super-Alfv{\'e}nic (${\mathcal M}_{\rm A} \sim 5$). Thus, except in the case where motions are transonic and weak shocks become important, motions are fundamentally hydrodynamic and turbulence follows a Kolmogorov ($s=5/3$) spectrum. However, while the turbulent energy density $U_{\rm t} \propto l^{2/3}$ decreases at small scales, the magnetic energy density is scale-independent. Thus, at some scale $l_{\rm A} \sim l_0 {\mathcal M}_{\rm A}^{-3}$, where $U_{\rm turb} \sim U_{B}$, the magnetic field becomes dynamically important, and turbulence transitions to the MHD regime. The cutoff scale in equation \ref{eqn:Wk} can be computed by setting the turbulent cascade rate for fast modes to the transit time damping rate on thermal electrons. This yields \citep{brunetti07,miniati15}:
\begin{equation}
k_{\rm c} \approx A k_{\rm outer} \frac{\langle \varv_{\rm outer}^{2}\rangle^{2}}{c_{\rm s}^{4}}
\label{eqn:k_c} 
\end{equation}
where $A \approx 11 000$. This gives $2\upi/k_\rmn{cut}\sim 0.1-1$~kpc in the ICM. This constitutes an
effective mean free path for CRs, unless plasma instabilities can
mediate interactions between turbulence and particles on smaller
scales \citep{brunetti11}, a possibility we discuss in \S\ref{sect:self-limiting}. Another possibility is that compressible modes dissipate in weak shocks, resulting in Burgers' turbulence $s=2$ \citep{kowal10,porter15,miniati15}. If Burgers turbulence dominates, then particle acceleration rates are too slow in the face of cooling processes to explain radio halos \citep{miniati15}, and an alternative model for radio halos is required.

@@

The important effects are best summarized in terms of the acceleration time: 
\begin{equation}
\tau_{\rm D}= \frac{p^{2}}{4 D_{\rm pp}}
\end{equation}
which is momentum independent. This should be compared against the lifetime of turbulence, $\tau_{\rm cl}$, and the cooling time $\tau_{\rm cool}$. In Table~\ref{tab:timescales}, we show both the thermal quantities and the
timescales for CR cooling and (re)acceleration for three different
spatial regions of the RH. The densities \citep{1992A&A...259L..31B} and temperatures \citep{2009ApJ...696.1886B,2001A&A...365L..67A} are derived from X-ray observations. To calculate synchrotron cooling times, we use B-fields derived from Faraday rotation measurements \citep{bonafede10}. To calculate the acceleration time, we need to assume an outer scale. We assume an injection scale $k_{\rm L}=2\pi/\lambda_{\rm L}$, where $\lambda_{\rm L} = 100$ kpc, which were assumptions adopted in previous work \citep{2006MNRAS.366.1437S,brunetti07, brunetti11}. This lengthscale corresponds to an eddy turnover time on the outer scale of $\varv_{L} k_{\rm L}^{-1} \sim 1.2$Gyr if $\varv_{\rm L} \sim 500 {\rm km \, s^{-1}}$, as is characteristic of a merger. Note that hydrodynamical simulations of clusters have sometimes found larger $\lambda_{\rm L}$, in some cases comparable to the size of the cluster \citep{vazza11,miniati15}. This choice is degenerate with $X_{\rm tu}$; in \S\ref{sect:self-limiting} we argue that $k_{\rm A}$ is a more appropriate choice, but also find that when the decay time is appropriately scaled, we are relatively insensitive to the choice of $k_{\rm L}$. We present $\tau_{\rm D}$ for 3 different models, which we discuss in the next section. The reacceleration timescale
$\tau_{\rm D}$ is similar between our three models, where the
difference comes from turbulent profile parameterized with
$\alpha_\rmn{tu}$. This implies that even small differences in the
turbulent profile could impact the seed CRs significantly. Finally, we adopt an duration of acceleration $\tau_{\rm cl} = 650$ Myr, in line with previous assumptions \citep{brunetti07}, roughly corresponding to the turbulent decay time. 
\begin{table}
  \caption{Thermal quantities and timescales for different spatial
    regions in a Coma like cluster.}
\begin{tabular}{l c  c c}
\hline
\hline
& & spatial regions & \\
 & $0.1\,\RH^{(2)}$ & $0.3\,\RH^{(2)}$ & $\RH^{(2)}$   \\
\hline
thermal quantities$^{(1)}$ & & & \\
$\rho$ [$10^{-27}$ g cm$^{-3}$] & 3.0 & 1.6 & 0.15 \\
$T$ [$10^{8}$ K] & 1.4 & 1.0 & 0.58 \\
\hline
timescales$^{(1)}$ & & & \\
$\tau_\rmn{D}$(\Mprimary)$^{(3)}$ [Gyr] & 0.45 & 0.44 & 0.39 \\
$\tau_\rmn{D}$(\Mstream)$^{(3)}$  [Gyr] & 0.50 & 0.47 & 0.34 \\
$\tau_\rmn{D}$(\Mflatturb)$^{(3)}$  [Gyr] & 0.69 & 0.56 & 0.27 \\
$\tau_\rmn{IC/sync}(P=10^4\,m_ec)^{(4)}$ [Gyr] & 0.11 & 0.15 & 0.22 \\
$\tau_\rmn{had}(P=100\,m_pc)^{(5)}$ [Gyr] & 2.4 & 4.5 & 47 \\
$\tau_\rmn{Coul}(P=\,m_ec)^{(6)}$  [Gyr] & 0.0092 & 0.017 & 0.17 \\
\hline
\end{tabular}
\begin{quote}
 Notes: \\ 
 (1) Median quantities from our simulated post-merging cluster g72a derived 
 during last 300~Myrs in time. \\
 (2) Radius of the giant radio halo in Coma where $\RH\approx0.6\,R_{200}$.\\
 (3) Fermi-II reacceleration for both electrons and protons at all energies.\\
 (4) Inverse Compton and synchrotron cooling for electrons.\\
 (5) Catastrophic losses for protons.\\
 (6) Coulomb cooling for electrons (protons factor $m_e/m_p$ smaller).\\
 \label{tab:timescales}
  \end{quote}
\end{table}

@@

We have seen that the radio luminosity depends very sensitively on $\tau_{\rm cl}/\tau_{\rm D}$, through the very sensitive dependence on $X_{\rm tu}$ (Fig \ref{fig:param_comp}; note from equation \ref{eq:Dpp_scaling} that $\tau_{\rm D} \propto X_{\rm tu}^{-2}$). Since this raises questions of fine-tuning in $X_{\rm tu}$ to explain the observations, it is worth understanding in more detail. Let us ignore cooling. This is a good approximation for the CR protons, for instance, since hadronic cooling times are long\footnote{In principle, CRp can still lose energy via wave heating, at a rate $\varv_{\rm A} \cdot \nabla P_{c}$, but we assume CR streaming is suppressed during mergers, when they are spatially confined by scattering.} In this case, $\dot{p}= p/\tau_{\rm D}$, and after a time $\tau_{\rm cl}$, we have $p \rightarrow p \, {\rm exp}(\tau_{\rm cl}/\tau_{\rm D})$ (where we have used the fact that $\tau{\rm D}$ is momentum independent). For an initial power law distribution function $f(p) = C p^{-\alpha}$, this momentum increase can be rewritten as a change of normalization, $f(p) = \tilde{C} p^{-\alpha}$, where $\tilde{C} = {\rm exp}( \alpha \tau_{\rm cl}/\tau_{\rm D})$. A slightly more careful derivation by direct solution of the Fokker-Planck equation yields: 
\begin{equation}
  \dot{f} = {{\partial }\over{\partial p}}
  p^2 D_{\rm pp} {{\partial }\over{\partial p}} \frac{f}{p^2}\,,
\end{equation}
with the analytic solution given by
\begin{equation}
  \label{eq:Creacc}
  \tilde{C}= 
  C \exp{\left[\frac{(2+\alpha)(\alpha-1)}{4}\frac{\tau_\rmn{cl}}{\tau_\rmn{D}}\right]}\,.
\end{equation}
This exponential sensitivity to $\tau_{\rm cl}/\tau_{\rm D}$ lies behind the extreme sensitivity to $X_{\rm tu}$ in Fig \ref{fig:param_comp}.

@@

An obvious candidate for $\tau_{\rm decay}$ is: 
\begin{equation}
\tau_{\rm edd} = \frac{L}{v_{\rm c}}
\end{equation}
i.e., the eddy turnover time at the outer scale. This is subject to uncertainties about the location of the outer scale $L$; estimates in the literature range from $L \sim 0.1-1$ Mpc. It is also worth remembering that MHD turbulence only applies below $l < l_{\rm A}$, where $l_{\rm A}$ is the Alfv{\'e}n scale where $\varv \sim \varv_{\rm A}$. Invoking fast modes, Kraichnan scalings, etc, is only valid below these scales. For $l > l_{\rm A}$, turbulence is basically hydrodynamic and Kolmogorov in nature. The momentum diffusion coefficient can thus be split up into two components: 
\begin{equation}
D_{\rm pp} = D_{\rm pp}^{\rm HD}[l_{\rm A} < l < L] + D_{\rm pp}^{\rm MHD}[l_{\rm c} < l < l_{\rm A}]
\end{equation}
From equations (\ref{eqn:diffusion}) and (\ref{eqn:k_W}), and assuming Kolmogorov turbulence from $l_{\rm A} < l < L$ and Kraichnan turbulence from $l_{\rm c} < l < l_{\rm A}$, we get: 
\begin{equation}
\frac{D_{\rm pp}^{\rm HD}}{D_{\rm pp}^{\rm MHD}} = \frac{k_{\rm L}^{2/3}k_{\rm A}^{1/3} \varv_{\rm c}^{2}}{k_{\rm A}^{1/2}k_{\rm c}^{1/2} \varv_{\rm A}^{2}} = A^{-1/2} \beta = 0.5 \left( \frac{\beta}{50} \right) 
\end{equation}
where we have used $k_{\rm A} = k_{\rm L} M_{A}^{3}$ (from Kolmogorov scalings, and $M_{\rm A} = \varv_{\rm c}/\varv_{\rm A}$ is the Alfv{\'e}n Mach number of the turbulence, $\beta \approx c_{\rm s}^{2}/\varv_{\rm A}^{2}$, and the cutoff scale $k_{\rm c}$ (equation \ref{eqn:k_c}).  Thus, the contribution of the two regimes are comparable (as we shall see later, this is not true if one incorrectly assumes Kraichnan scalings apply from the outer scale $L$; this boosts the contribution of the scales where $l_{\rm A} < l < L$ by a factor $M_{\rm A}^{1/2}$). 
Henceforth, we can consider the outer scale of the fast modes to be $l_{\rm A}$, with velocity $\varv_{\rm A}$. The turbulent reacceleration time is: 
\begin{equation}
\tau_{\rm D} = \frac{p^{2}}{4 D_{\rm pp}} = C_{\rm D} A^{1/2} c k_{\rm A} \frac{\varv_{A}^{2}}{c_{\rm S}^{2}}
\label{eqn:tau_D} 
\end{equation}
where $C_{\rm D} = 2/(5 \pi)$, $A\approx 11000$, and we have used equation \ref{eqn:k_c}. The turbulence decay time is given by the cascade time of fast modes (\citet{2004ApJ...614..757Y}; see equation \ref{eqn:cascade}): 
\begin{equation}
\tau_{\rm decay} = \frac{c_{\rm s}}{\varv_{\rm A}^{2} k_{\rm A}}\,.
\label{eqn:tau_decay} 
\end{equation} 
This can be related to the eddy turnover time at the outer scale $\tau_{\rm edd}$ from $l_{\rm A} = L M_{\rm A}^{-3}$ to: 
\begin{equation}
\tau_{\rm decay} =\frac{L}{\varv_{\rm c}} \left(\frac{c_{\rm s}}{\varv_{\rm c}} \right)^{2} \frac{\varv_{\rm A}}{c_{\rm s}} = 0.7 \, \tau_{\rm edd} \left(\frac{\tilde{X}_{\rm tu}}{0.2}\right)^{-1}  \left(\frac{\beta}{100} \right)^{-1/2}
\end{equation}
where we have used $\tilde{X}_{\rm tu} \approx (\varv_{\rm c}/c_{\rm s})^{2}$ and $c_{\rm s}/\varv_{\rm A} = \beta^{1/2}$. Thus $\tau_{\rm decay} \sim \tau_{\rm edd}$, but we prefer to use the expression for $\tau_{\rm edd}$, since it is more physically correct and leads to more revealing scalings.

From equations \ref{eqn:tau_D} and \ref{eqn:tau_decay}, we obtain:
\begin{equation}
\frac{\tau_{\rm cl}}{\tau_{\rm D}} \approx \frac{\tau_{\rm decay}}{\tau_{\rm D}} = 0.1 \left( \frac{c_{\rm s}}{1500 \, {\rm km \, s^{-1}}} \right) \left( \frac{\beta}{50} \right)^{-1}\,. 
\label{eqn:ratios_TTD} 
\end{equation}
Remarkably, this expression is independent of properties of the turbulence such as $\varv_{\rm c},L$, or $l_{\rm A}$, and depends only on properties of the plasma ($c_{\rm s},\beta$). Ultimately, this arises because the timescale on which a fast mode wave transfers energy due to wave particle interactions in second order Fermi acceleration, $\tau_{\rm p} \sim (k_{\rm p} c (\varv_{\rm c}/c)^{2} )^{-1}$, is closely related to the timescale on which it cascades due to wave-wave interactions, $\tau_{\rm w} \sim c_{\rm s}/(k_{\rm w} \varv_{\rm c}^{2})$. This implies $\tau_{\rm w}/\tau_{\rm p} \sim (c_{s}/c)(l_{\rm w}/l_{\rm p})$, where $l_{\rm w} \sim l_{\rm A}$ is the outer scale on which the fast mode cascade begins, and $l_{\rm p} \sim (l_{\rm A} l_{\rm c})^{1/2}$ is the characteristic wavelength for wave-particle interactions. For transit time damping and an outer scale of $l_{\rm A}$, we have $l_{\rm c} \sim (m_{\rm e}/m_{\rm p}) l_{\rm A} (c_{\rm s}/\varv_{\rm A})^{4}$. We thus have $\tau_{\rm w}/\tau_{\rm p} \sim (c_{\rm s}/c)(l_{\rm A}/l_{\rm c})^{1/2} \sim (c_{\rm s}/c) (m_{\rm p}/m_{\rm e})^{1/2} \beta^{-1}$. More careful consideration of dimensionless factors boosts this estimate by an order of magnitude to give equation \ref{eqn:ratios_TTD}.
